---
title: "Final Project"
author: "Rakel Faaland"
date: "2022-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

if (!require('tidyverse')) 
  install.packages('tidyverse', 
                   repos = "http://cran.us.r-project.org"); 
library('tidyverse')

if (!require('supernova')) 
  install.packages('supernova', 
                   repos = "http://cran.us.r-project.org"); 
library('supernova')

if (!require('kableExtra')) 
  install.packages('kableExtra', 
                   repos = "http://cran.us.r-project.org"); 
library('kableExtra')

```

```{r data_import}

# NYC Population 

nyc_population <- jsonlite::fromJSON("https://data.cityofnewyork.us/resource/xywu-7bv9.json?$limit=40")

# NYC Poverty Metrics

evictions <- jsonlite::fromJSON("https://data.cityofnewyork.us/resource/6z8x-wfk4.json?$limit=100000")

snap_recipients <- jsonlite::fromJSON("https://data.cityofnewyork.us/resource/jye8-w4d7.json?$limit=2000")

# NYC Resources

homebases <- jsonlite::fromJSON("https://data.cityofnewyork.us/resource/ntcm-2w4k.json?$limit=200")

shelters <- jsonlite::fromJSON("https://data.cityofnewyork.us/resource/3qem-6v3v.json?$limit=8000")

employment_programs <- 
jsonlite::fromJSON("https://data.cityofnewyork.us/resource/99br-frp6.json?$limit=400")

financial_centers <- 
jsonlite::fromJSON("https://data.cityofnewyork.us/resource/dt2z-amuf.json?$limit=40")

community_centers <-
jsonlite::fromJSON("https://data.cityofnewyork.us/resource/crns-fw6u.json?$limit=14000")

snap_centers <- 
jsonlite::fromJSON("https://data.cityofnewyork.us/resource/tc6u-8rnp.json?$limit=40")

```

```{r cleaning}

# cleaning nyc_population

# selecting only the variables on 2020
# renaming colums to represent the "count" and "percent"
# subsetting the rows to only include the 5 boroughs, NOT the NYC total

nyc_population_clean <- nyc_population%>%
  select(borough, "_2020", "_2020_boro_share_of_nyc_total")%>%
  rename("population_count" = "_2020", "population_percent" = "_2020_boro_share_of_nyc_total")%>%
  mutate(borough = toupper(borough))%>%
  mutate(population_count = as.numeric(population_count))%>%
  mutate(population_percent = as.numeric(population_percent))

# SOURCE (subsetting rows): https://www.r-bloggers.com/2020/08/how-to-subset-rows-from-a-data-frame-in-r/

nyc_population_clean <- nyc_population_clean[c(2:6),]

# cleaning evictions

# selecting the variables for the type of building (residential or commercial) and the borough
# filtering for only residential building evictions

evictions_clean <- evictions%>%
  select(residential_commercial_ind, borough)%>%
  rename("building_type" = residential_commercial_ind)%>%
  filter(building_type == "Residential")

 # Negative indexing the "building_type" variable because I don't need it anymore (I already filtered for "Residential" only)

evictions_clean <- evictions_clean[-c(1)]

# Calculating the number of evictions per borough using count function
# Calculating the percent of evictions per borough

evictions_clean <- evictions_clean%>%
  count(borough)%>%
  rename("evictions_count" = n)%>%
  mutate(evictions_percent = round(evictions_count/sum(evictions_count)*100))

# cleaning snap_recipients

snap_recipients_clean <- snap_recipients%>%
  select(boro, bc_snap_recipients)%>%
  rename("borough" = boro)%>%
  mutate(borough = toupper(borough))%>%
  mutate(bc_snap_recipients = as.numeric(bc_snap_recipients))%>%
  group_by(borough)%>%
  summarise(snap_recipients_count = sum(bc_snap_recipients))%>%
  mutate(snap_recipients_percent = round(snap_recipients_count/sum(snap_recipients_count)*100))

# cleaning homebases

homebases_clean <- homebases%>%
  select(homebase_office,
         service_area_zip_code, borough, postcode)%>%
  rename("servicing_zipcodes" = service_area_zip_code,
         "zipcode" = postcode)%>%
  mutate(homebase_office = toupper(homebase_office))%>%
  mutate(borough = ifelse(borough == "STATEN IS", "STATEN ISLAND", borough))%>%
  mutate(servicing_zipcodes =
            ifelse(servicing_zipcodes %in%
                     c("Queens", "Manhattan", "Staten Island"),
                   zipcode, servicing_zipcodes))%>%
  mutate(servicing_zipcodes_number =
           lengths(str_split(servicing_zipcodes, ",")))%>%
  count(borough)%>%
  rename("homebases_count" = n)%>%
  mutate(homebases_percent = round(homebases_count/sum(homebases_count)*100))

# More info on the homebases: https://www.nyc.gov/site/hra/help/homebase-locations.page

# cleaning shelters

shelters_clean <- shelters%>%
  select(borough)%>%
  mutate(borough = toupper(borough))%>%
  count(borough)%>%
  rename("shelters_count" = n)%>%
  mutate(shelters_percent =
           round(shelters_count/sum(shelters_count)*100))

# cleaning employment_programs

employment_programs_clean <- employment_programs%>%
  select(program, borough_community)%>%
  rename("borough" = borough_community)%>%
  mutate(borough = toupper(borough))%>%
  mutate(borough = ifelse(borough %in% "NEW YORK", "MANHATTAN", borough))%>%
  mutate(borough = str_squish(borough))%>%
  mutate(borough = ifelse(borough %in% "LONG ISLAND CITY", NA, borough))%>%
  drop_na()%>%
  mutate(program_type = ifelse(program %in% "Adult Employment", "Adult", "Youth"))

employment_programs_clean <- employment_programs_clean[-c(1)]

# cleaning financial_centers 

financial_centers_clean <- financial_centers%>%
  select(borough)%>%
  mutate(borough = toupper(borough))%>%
  count(borough)%>%
  rename("financial_centers_count" = n)%>%
  mutate(financial_centers_percent = round(financial_centers_count/sum(financial_centers_count)*100))

# cleaning community_centers

community_centers_clean <- community_centers%>%
  select(borough)%>%
  mutate(borough = toupper(borough))%>%
  drop_na()%>%
  count(borough)%>%
  rename("community_centers_count" = n)%>%
  mutate(community_centers_percent = round(community_centers_count/sum(community_centers_count)*100))

# cleaning snap_centers

snap_centers_clean <- snap_centers%>%
  select(borough)%>%
  mutate(borough = toupper(borough))%>%
  count(borough)%>%
  rename("snap_centers_count" = n)%>%
  mutate(snap_centers_percent = round(snap_centers_count/sum(snap_centers_count)*100))

# Removing the dirty data from my global environment:

# To remove things from the global env is rm(object_name)

rm(snap_centers, financial_centers, community_centers, employment_programs, evictions, homebases, shelters, snap_recipients, nyc_population)

```

``` {r extra_tables}

# Table of employment programs by borough

employment_programs_borough <- employment_programs_clean%>%
  count(borough)%>%
  rename("employment_programs_count" = n)%>%
  mutate(employment_programs_percent = round(employment_programs_count/sum(employment_programs_count)*100))

# Table of adult v. youth employment programs

employment_programs_type <- employment_programs_clean%>%
  count(program_type)%>%
  rename("count" = n)%>%
  mutate(percent = round(count/sum(count)*100))

# Removing employment_programs_clean because I don't need it anymore 

rm(employment_programs_clean)

```

```{r merging, include = FALSE}

# Merging all nyc poverty indicator data frames

poverty_indicators_merge <- cbind(evictions_clean, snap_recipients_clean)

# Deleting extra "borough"columns

colnames(poverty_indicators_merge)

poverty_indicators_merge <- poverty_indicators_merge[-c(4)]
  
# POVERTY INDICATOR (count) by BOROUGH (for future table)

# Delete all "percent" columns

borough_indicator_count <- poverty_indicators_merge[-c(3,5)]
  

# POVERTY INDICATOR (percent) by BOROUGH (for future bar graph)

# Delete all "count" columns

borough_indicator_percent <- poverty_indicators_merge[-c(2,4)]

# Merging all nyc resource related data frames

resources_merge <- cbind(homebases_clean, shelters_clean, employment_programs_borough, financial_centers_clean, community_centers_clean, snap_centers_clean)

# RESOURCES (count) BY BOROUGH (for future table)

# Delete all extra "borough" columns + all "percent" columns

colnames(resources_merge)

borough_resources_count <- resources_merge[-c(3,4,6,7,9,10,12,13,15,16,18)]

# Creating dummy variables for each borough (for creating models)
borough_resources_count <- borough_resources_count%>% 
  mutate(borough_bronx = ifelse(borough %in% c("BRONX"), 1,0), 
         borough_brooklyn = ifelse(borough %in% c("BROOKLYN"), 1, 0 ), 
         borough_queens = ifelse(borough %in% c("QUEENS"), 1,0), 
         borough_manhattan = ifelse(borough %in% c("MANHATTAN"), 1, 0 ), 
         borough_staten_island = ifelse(borough %in% c("STATEN ISLAND"), 1,0)
         )

# RESOURCES (percent) BY BOROUGH (for future bar graph)

# Delete all extra "borough" columns + all "count" columns

borough_resources_percent <- resources_merge[-c(2,4,5,7,8,10,11,13,14,16,17)]

# Removing unnessary objects from global environment:

rm(poverty_indicators_merge, resources_merge)

```

```{r borough_resources_stats, include = FALSE}

homebase_stats <- homebases_clean%>%
  summarise(Minimum = min(homebases_count),
            Maximum = max(homebases_count),
            Mean = mean(homebases_count), 
            "Standard Deviation" = sd(homebases_count),
            Median = median(homebases_count),
            IQR = IQR(homebases_count))
  
shelter_stats <- shelters_clean%>%
  summarise(Minimum = min(shelters_count),
            Maximum = max(shelters_count),
            Mean = mean(shelters_count), 
            "Standard Deviation" = sd(shelters_count),
            Median = median(shelters_count),
            IQR = IQR(shelters_count))

financial_center_stats <- financial_centers_clean%>%
  summarise(Minimum = min(financial_centers_count),
            Maximum = max(financial_centers_count),
            Mean = mean(financial_centers_count), 
            "Standard Deviation" = sd(financial_centers_count),
            Median = median(financial_centers_count),
            IQR = IQR(financial_centers_count))
  
community_center_stats <- community_centers_clean%>%
  summarise(Minimum = min(community_centers_count),
            Maximum = max(community_centers_count),
            Mean = mean(community_centers_count), 
            "Standard Deviation" = sd(community_centers_count),
            Median = median(community_centers_count),
            IQR = IQR(community_centers_count))

employment_program_stats <- employment_programs_borough%>%
  summarise(Minimum = min(employment_programs_count),
            Maximum = max(employment_programs_count),
            Mean = mean(employment_programs_count), 
            "Standard Deviation" = sd(employment_programs_count),
            Median = median(employment_programs_count),
            IQR = IQR(employment_programs_count))

snap_center_stats <- snap_centers_clean%>%
  summarise(Minimum = min(snap_centers_count),
            Maximum = max(snap_centers_count),
            Mean = mean(snap_centers_count), 
            "Standard Deviation" = sd(snap_centers_count),
            Median = median(snap_centers_count),
            IQR = IQR(snap_centers_count))

borough_resource_stats <- rbind(homebase_stats, shelter_stats, financial_center_stats, community_center_stats, employment_program_stats, snap_center_stats)

# SOURCE (re-ordering variables): http://www.sthda.com/english/wiki/reordering-data-frame-columns-in-r

# Creating a vector with the names of each resource (to add to the resource_stats data frame as the first column)

resource = c("Homebase Offices", "Shelters", "Financial Empowerment Centers", "Communities Center", "Employment Programs", "SNAP Centers")

# Saving resource as a new column in resource_stats

borough_resource_stats$Resource <- resource

# Finding the column names and their index number

colnames(borough_resource_stats)

# re-ordering the columns by index number so that "resources" (7) is first

borough_resource_stats <- borough_resource_stats[, c(7,1,2,3,4,5,6)]

# Deleting unnessary objects from envirionment:

rm(homebase_stats, shelter_stats, employment_program_stats, financial_center_stats, community_center_stats, snap_center_stats)

```

```{r aesthetic_tables}

# Creating new objects so I can apply kableExtra functions for aesthetics

# (1) borough_resources_count_table

# Delete dummy variables

borough_resources_count_table <- borough_resources_count[-c(8,9,10,11,12)]
  
borough_resources_count_table <- borough_resources_count_table%>%
  rename("Borough" = borough, "Hombases" = homebases_count, "Shelters" = shelters_count, "Employment Programs" = employment_programs_count, "Financial Empowerment Centers" = financial_centers_count, "Community Centers" = community_centers_count, "SNAP Centers" = snap_centers_count)%>%
  kbl(caption = "NYC Resources by Borough")%>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "hover"))

# (2) borough_resource_stats_table

borough_resource_stats_table <- borough_resource_stats%>%
  mutate(`Standard Deviation` = round(`Standard Deviation`))%>%
  kbl(caption = "Statistics on NYC Resources by Borough")%>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "hover"))

# (3) bronx_resources_table

bronx_resources_table <- borough_resources_count%>%
  select(borough, homebases_count, shelters_count, employment_programs_count, financial_centers_count, community_centers_count, snap_centers_count)%>%
  rename("Borough" = borough, "Hombases" = homebases_count, "Shelters" = shelters_count, "Employment Programs" = employment_programs_count, "Financial Empowerment Centers" = financial_centers_count, "Community Centers" = community_centers_count, "SNAP Centers" = snap_centers_count)%>%
  filter(Borough %in% "BRONX")%>%
  kbl(caption = "NYC Resources in the Bronx")%>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "hover"))

# (4) brooklyn_resources_table

brooklyn_resources_table <- borough_resources_count%>%
  select(borough, homebases_count, shelters_count, employment_programs_count, financial_centers_count, community_centers_count, snap_centers_count)%>%
  rename("Borough" = borough, "Hombases" = homebases_count, "Shelters" = shelters_count, "Employment Programs" = employment_programs_count, "Financial Empowerment Centers" = financial_centers_count, "Community Centers" = community_centers_count, "SNAP Centers" = snap_centers_count)%>%
  filter(Borough %in% "BROOKLYN")%>%
  kbl(caption = "NYC Resources in Brooklyn")%>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "hover"))

# (5) manhattan_resources_table

manhattan_resources_table <- borough_resources_count%>%
  select(borough, homebases_count, shelters_count, employment_programs_count, financial_centers_count, community_centers_count, snap_centers_count)%>%
  rename("Borough" = borough, "Hombases" = homebases_count, "Shelters" = shelters_count, "Employment Programs" = employment_programs_count, "Financial Empowerment Centers" = financial_centers_count, "Community Centers" = community_centers_count, "SNAP Centers" = snap_centers_count)%>%
  filter(Borough %in% "MANHATTAN")%>%
  kbl(caption = "NYC Resources in Manhattan")%>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "hover"))

# (6) queens_resources_table

queens_resources_table <- borough_resources_count%>%
  select(borough, homebases_count, shelters_count, employment_programs_count, financial_centers_count, community_centers_count, snap_centers_count)%>%
  rename("Borough" = borough, "Hombases" = homebases_count, "Shelters" = shelters_count, "Employment Programs" = employment_programs_count, "Financial Empowerment Centers" = financial_centers_count, "Community Centers" = community_centers_count, "SNAP Centers" = snap_centers_count)%>%
  filter(Borough %in% "QUEENS")%>%
  kbl(caption = "NYC Resources in Queens")%>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "hover"))

# (7) staten_island_resources_table

staten_island_resources_table <- borough_resources_count%>%
  select(borough, homebases_count, shelters_count, employment_programs_count, financial_centers_count, community_centers_count, snap_centers_count)%>%
  rename("Borough" = borough, "Hombases" = homebases_count, "Shelters" = shelters_count, "Employment Programs" = employment_programs_count, "Financial Empowerment Centers" = financial_centers_count, "Community Centers" = community_centers_count, "SNAP Centers" = snap_centers_count)%>%
  filter(Borough %in% "STATEN ISLAND")%>%
  kbl(caption = "NYC Resources in Staten Island")%>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "hover"))

# (8) borough_indicator_table

borough_indicator_table <- borough_indicator_count%>%
  rename("Borough" = borough, "Residential Evictions" = evictions_count, "SNAP Recipients" = snap_recipients_count)%>%
   kbl(caption = "NYC Poverty Indicators")%>%
  kable_styling(bootstrap_options = c("bordered", "condensed", "hover"))

rm(borough_indicator_count)

```

```{r visualizations}

# (1) population_bar

v2_population_percent <- nyc_population_clean%>%
  select(borough, population_percent)%>%
  rename("Borough" = borough, "Percent" = population_percent)
 
population_bar <- v2_population_percent%>%
  ggplot(
    aes(
      x = Borough,
      y = Percent,
      fill = Borough)) + 
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = "black"
  ) +
  labs(
    title = "NYC Population by Borough",
    subtitle = "Borough Indicated by Color",
    caption = "Data from NYC OPEN DATA")+ 
  ylab("Population (Percent)")

# (2) borough_resources_percent_bar

# Renaming variables to remove "_" and capitalize first letter

v2_borough_resources_percent <- borough_resources_percent%>%
  rename("Borough" = borough, "Hombases" = homebases_percent, "Shelters" = shelters_percent, "Employment Programs" = employment_programs_percent, "Financial Empowerment Centers" = financial_centers_percent, "Community Centers" = community_centers_percent, "SNAP Centers" = snap_centers_percent)

# Changing the format of the data to be able to plot multiple variables (all resources)

# SOURCE: https://stackoverflow.com/questions/46916042/simple-bar-plot-with-multiple-variables-in-r-similar-to-excel

long.borough_resources_percent <- gather(v2_borough_resources_percent, variable, value, -Borough)%>%
  rename("Percent" = value, "Resources" = variable)
  
# Creating single bar graph 

borough_resources_percent_bar <- ggplot(long.borough_resources_percent,
    aes(
      x = Borough,
      y = Percent,
      fill = Resources)) + 
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = "black"
  ) +
  labs(
    title = "Percentage of NYC Resources by Borough",
    subtitle = "Resource Type Indicated by Color",
    caption = "Data from NYC OPEN DATA")

# Creating detailed bar graph (facet_wrap)

borough_resources_percent_bar_detailed <- ggplot(long.borough_resources_percent,
    aes(
      x = Borough,
      y = Percent,
      fill = Resources)) + 
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = "black"
  ) +
  coord_flip() +
  labs(
    title = "Percentage of NYC Resources by Borough",
    subtitle = "Resource Type Indicated by Color",
    caption = "Data from NYC OPEN DATA") +
  facet_wrap(~Resources) + 
  theme(legend.position = "none")

# (2) bronx_resources_bar

long.bronx_resources <- long.borough_resources_percent%>%
  filter(Borough %in% "BRONX")
  
bronx_resources_bar <- ggplot(long.bronx_resources,
    aes(
      x = Borough,
      y = Percent,
      fill = Resources)) + 
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = "black"
  ) +
  labs(
    title = "Percentage of NYC Resources in the Bronx",
    subtitle = "Resource Type Indicated by Color",
    caption = "Data from NYC OPEN DATA")+
  geom_label(aes(x=.625,y=19, label= "21%"), fill = "white") +
  geom_label(aes(x=.775,y=20, label= "22%"), fill = "white") +
  geom_label(aes(x=.925,y=15, label= "17%"), fill = "white") +
  geom_label(aes(x=1.075,y=36, label= "38%"), fill = "white") +
  geom_label(aes(x=1.225,y=21, label= "23%"), fill = "white") +
  geom_label(aes(x=1.375,y=19, label= "21%"), fill = "white")

# (3) brooklyn_resources_bar

long.brooklyn_resources <- long.borough_resources_percent%>%
  filter(Borough %in% "BROOKLYN")
  
brooklyn_resources_bar <- ggplot(long.brooklyn_resources,
    aes(
      x = Borough,
      y = Percent,
      fill = Resources)) + 
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = "black"
  ) +
  labs(
    title = "Percentage of NYC Resources in the Brooklyn",
    subtitle = "Resource Type Indicated by Color",
    caption = "Data from NYC OPEN DATA")+
  geom_label(aes(x=.625,y=27, label= "29%"), fill = "white") +
  geom_label(aes(x=.775,y=34, label= "36%"), fill = "white") +
  geom_label(aes(x=.925,y=24, label= "26%"), fill = "white") +
  geom_label(aes(x=1.075,y=36, label= "38%"), fill = "white") +
  geom_label(aes(x=1.225,y=28, label= "30%"), fill = "white") +
  geom_label(aes(x=1.375,y=19, label= "21%"), fill = "white")
  
# (4) manhattan_resources_bar

long.manhattan_resources <- long.borough_resources_percent%>%
  filter(Borough %in% "MANHATTAN")
  
manhattan_resources_bar <- ggplot(long.manhattan_resources,
    aes(
      x = Borough,
      y = Percent,
      fill = Resources)) + 
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = "black"
  ) +
  labs(
    title = "Percentage of NYC Resources in the Manhattan",
    subtitle = "Resource Type Indicated by Color",
    caption = "Data from NYC OPEN DATA")+
   geom_label(aes(x=.625,y=34, label= "36%"), fill = "white") +
  geom_label(aes(x=.775,y=18, label= "20%"), fill = "white") +
  geom_label(aes(x=.925,y=33, label= "35%"), fill = "white") +
  geom_label(aes(x=1.075,y=6, label= "8%"), fill = "white") +
  geom_label(aes(x=1.225,y=20, label= "22%"), fill = "white") +
  geom_label(aes(x=1.375,y=27.2, label= "29%"), fill = "white")

# (5) queens_resources_bar

long.queens_resources <- long.borough_resources_percent%>%
  filter(Borough %in% "QUEENS")
  
queens_resources_bar <- ggplot(long.queens_resources,
    aes(
      x = Borough,
      y = Percent,
      fill = Resources)) + 
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = "black"
  ) +
  labs(
    title = "Percentage of NYC Resources in the Queens",
    subtitle = "Resource Type Indicated by Color",
    caption = "Data from NYC OPEN DATA")+
  geom_label(aes(x=.625,y=8.7, label= "10%"), fill = "white") +
  geom_label(aes(x=.775,y=18.7, label= "20%"), fill = "white") +
  geom_label(aes(x=.925,y=15.7, label= "17%"), fill = "white") +
  geom_label(aes(x=1.075,y=6.7, label= "8%"), fill = "white") +
  geom_label(aes(x=1.225,y=21.7, label= "23%"), fill = "white") +
  geom_label(aes(x=1.375,y=19.56, label= "21%"), fill = "white")

# (6) staten_island_resources_bar
  
long.staten_island_resources <- long.borough_resources_percent%>%
  filter(Borough %in% "STATEN ISLAND")
  
staten_island_resources_bar <- ggplot(long.staten_island_resources,
    aes(
      x = Borough,
      y = Percent,
      fill = Resources)) + 
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = "black"
  ) +
  labs(
    title = "Percentage of NYC Resources in the Staten Island",
    subtitle = "Resource Type Indicated by Color",
    caption = "Data from NYC OPEN DATA")+
  geom_label(aes(x=.625,y=4.6, label= "5%"), fill = "white") +
  geom_label(aes(x=.775,y=1.6, label= "2%"), fill = "white") +
  geom_label(aes(x=.925,y=3.6, label= "4%"), fill = "white") +
  geom_label(aes(x=1.075,y=7.6, label= "8%"), fill = "white") +
  geom_label(aes(x=1.225,y=1.6, label= "2%"), fill = "white") +
  geom_label(aes(x=1.375,y=6.56, label= "7%"), fill = "white")

# (7) borough_indicator_bar

v2_borough_indicator_percent <- borough_indicator_percent%>%
  rename("Borough" = borough, "Residential Evictions" = evictions_percent, "SNAP Recipients" = snap_recipients_percent)
  
long.borough_indicator_percent <- gather(v2_borough_indicator_percent, variable, value, -Borough)%>%
  rename("Percent" = value, "Indicators" = variable)
  
# Creating bar graph 

borough_indicator_bar <- ggplot(long.borough_indicator_percent,
    aes(
      x = Borough,
      y = Percent,
      fill = Indicators)) + 
  geom_bar(
    stat = "identity",
    position = "dodge",
    color = "black"
  ) +
  labs(
    title = "NYC Poverty Indicators by Borough",
    subtitle = "Indicator Specified by Color",
    caption = "Data from NYC OPEN DATA")

# Removing unnessary objects

rm(v2_population_percent,
   v2_borough_resources_percent,
   v2_borough_indicator_percent, 
   long.borough_resources_percent,
   long.borough_indicator_percent,
   long.staten_island_resources,
   long.queens_resources,
   long.manhattan_resources,
   long.brooklyn_resources,
   long.bronx_resources)

```

   
# Analysis of NYC Resources by Borough{.tabset}

## **INTRODUCTION**

### Research Question:

I am interested in exploring the following research question: **What resources are available to New Yorkers living in poverty or at risk of homelessness?**

In New York City, there are a lot of people experiencing homelessness, and the problem has only increased since the COVID-19 pandemic. According to predictions by the United States Census Bureau, 17% of New York City live in poverty. Solving the issue of poverty and homelessness is complex, and every individual has a unique set of needs that requires different services. In my analysis, I want to investigate what resources are available to serve those living in poverty and those at risk of, or experiencing, homelessness.
  
### Data Sources: 

For this project, I use NYC Open Data as the source of all my data. NYC Open Data provides free data to the public from various New York City government agencies. The NYC Office of Technology and Innovation manages NYC Open Data. The datasets I will be using are detailed below.
  
- [“Directory Of Homebase Locations”](https://data.cityofnewyork.us/Social-Services/Directory-Of-Homebase-Locations/ntcm-2w4k)

    + The “Directory Of Homebase Locations” dataset is provided by the Department of Homeless Services (DHS). The DHS is a government organization that provides services to New Yorkers who are experiencing homelessness. 

    + This dataset has 26 rows and 14 columns. Each observation is a homebase office. The variables in this dataset are the following: Provider, Service Area Zipcode, Address, Phone Number, Postcode, Borough, Latitude, Longitude, Community Board, Council District, Census Tract, BIN, BBL, NTA.

- [“Evictions”](https://data.cityofnewyork.us/City-Government/Evictions/6z8x-wfk4)

    + The “Evictions” dataset is provided by the Department of Investigation (DOI). The DOI is a government law enforcement agency that conducts covert investigations within the government and in organizations receiving benefits from the New York City government programs. 

    + This dataset has 71.2K rows and 20 columns. Each observation is an eviction. The variables in this dataset are the following: Court Index Number, Docket Number, Eviction Address, Evictions Apartment Number, Executed Date, Marshal First Name, Marshal Last Name, Residential/Commercial, Borough, Eviction Postcode, Ejectment, Eviction/Legal Possession, Latitude, Longitude, Community Board, Council District, Census Tract, BIN, BBL, NTA. 

- [“Buildings by Borough and Community District”](https://data.cityofnewyork.us/Social-Services/Buildings-by-Borough-and-Community-District/3qem-6v3v)

    + The “Buildings by Borough and Community District” dataset is provided by the Directory of Homeless Services (DHS). The DHS is a government organization that provides services to New Yorkers who are experiencing homelessness.
    
    + This dataset has 2,578 rows and 10 columns. Each observation is a shelter. The variables in this dataset are the following: Report Date, Borough, Community District, Adult Family Community Hotel, Adult Family Shelter, Adult Shelter, Adult Shelter Community Hotel, Families With Children Cluster, Families With Children Community Hotel, Families With Children Shelter.

- [“Borough/Community District Report - SNAP Population”](https://data.cityofnewyork.us/Social-Services/Borough-Community-District-Report-SNAP-Population/jye8-w4d7)

    + The “Borough/Community District Report - SNAP Population” dataset is provided by the Human Resources Administration (HRA). The HRA is a government organization that helps low-income New Yorkers with food assistance and emergency rental assistance. 
    
    + This dataset has 1,003 rows and 5 columns. Each observation is a monthly report on a community district. The variables in this dataset are the following: Month, Borough, Community District Number, SNAP recipients, SNAP households. 

- [“DYCD after-school programs: Jobs and Internships”](https://data.cityofnewyork.us/Education/DYCD-after-school-programs-Jobs-and-Internships/99br-frp6)

    + The “DYCD after-school programs: Jobs and Internships” dataset is provided by the Department of Youth and Community Development (DYCD). The DYCD is a government organization that supports youth and families in New York by funding community organizations. 
    
    + This dataset has 208 rows and 18 columns. Each observation is a program. The variables in this dataset are the following: Program Type, Program, Site Name, Address, Borough, Postcode, Agency, Contact Number, Grade Level / Age Group, Location 1, Latitude, Longitude, Community Board, Community Council, Census Tract, BIN, BBL, NTA.

- [“Financial Empowerment Centers”](https://data.cityofnewyork.us/Business/Financial-Empowerment-Centers/dt2z-amuf)

    + The “Financial Empowerment Centers” dataset is provided by the Department of Consumer and Worker Protection (DCWP). The DCWP is a government organization that supports businesses and individuals regarding finance. They license businesses, provide business resources, advocate for wages, and provide community outreach like their Financial Empowerment Centers.
    
    + This dataset has 23 rows and 23 columns. Each observation is a center. The variables in this dataset are the following: Provider, Host Organization, Building, Street, City, Borough, Postcode, Site Location Address, Days Open, Hours, Language(s), Telephone, Council District, Community Board, BBL, NTA, X coordinates, Y coordinates, Latitude, Longitude, Census Tract, BIN, Location. 

- [“Directory of NYCHA Community Facilities”](https://data.cityofnewyork.us/Social-Services/Directory-of-NYCHA-Community-Facilities/crns-fw6u)

    + The “Directory of NYCHA Community Facilities” dataset is provided by the New York City Housing Authority (NYCHA). NYCHA is a government organization that provides public housing to low-income New Yorkers. It also connects its recipients with opportunities related to career advancement, continuing education, and learning about financial health. 
    
    + This dataset has 430 rows and 16 columns. Each observation is a NYCHA community facility. The variables in this dataset are the following: Type, Funding Agency, Borough, Program Type, Status, Development, Address, Sponsor, Latitude, Longitude, Community Board, Council District, Census Tract, BIN, BBL, NTA.

- [“Directory of SNAP Centers”](https://data.cityofnewyork.us/Social-Services/Directory-of-SNAP-Centers/tc6u-8rnp)

    + The “Directory of SNAP Centers” dataset is provided by the Human Resources Administration (HRA). The HRA is a government organization that helps low-income New Yorkers with food assistance and emergency rental assistance. 
    
    + This dataset has 14 rows and 16 columns. Each observation is a SNAP center. The variables in this dataset are the following: Borough, Facility Name, Street Address, City, Postcode, State, Phone Number(s), Comments, Latitude, Longitude, Community Board, Council District, Census Tract, BIN, BBL, NTA.

- [“New York City Population by Borough, 1950 - 2040”](https://data.cityofnewyork.us/City-Government/New-York-City-Population-by-Borough-1950-2040/xywu-7bv9)

    + The “New York City Population by Borough, 1950 - 2040” dataset is provided by the Department of City Planning (DCP). The DCP is a government organization that oversees city land use for housing, neighborhood development, sustainability, and economic development. 
    
    + This dataset has 6 rows and 22 columns. Each observation represents a total. The 6 rows contain the 5 boroughs and a row for New York City as a whole. The variables in this dataset are the following: Age Group, Borough, 1950, 1950 - Borough Share of NYC Total, 1960, 1960 - Borough Share of NYC Total, 1970, 1970 - Borough Share of NYC Total,  1980, 1980 - Borough Share of NYC Total, 1990, 1990 - Borough Share of NYC Total, 2000, 2000 - Borough Share of NYC Total,  2010, 2010 - Borough Share of NYC Total, 2020, 2020 - Borough Share of NYC Total, 2030, 2030 - Borough Share of NYC Total, 2040, 2040 - Borough Share of NYC Total.

### Method:

In order to investigate my research question, I will use the data sets I have selected above to create three new data frames. These data frames will contain counts and percentages by the borough of the New York City population, New York City resources, and New York City poverty indicators. 

For the New York City population data frame, I will be using only the data from 2020. This data will give context to my later analysis of the distribution of resources and poverty indicators by depicting how New York City's population is spread throughout the five boroughs. 

The New York City resources data frame will contain the following variables: Homebase Offices, Shelters, Employment Programs, Financial Empowerment Centers, Community Centers, and SNAP Centers. This will show how these resources are distributed through the five boroughs of New York City.

The poverty indicators data frame will contain the following variables: Residential Evictions, and SNAP Recipients. These variables will be used to identify cases of people experiencing economic hardship and at risk of homelessness. It will help to identify where the greatest need for resources is in New York City's five boroughs.

In my analysis, I will compare the distribution of resources, poverty indicators, and population within each borough to provide a picture of the need and resource allocation in New York City by borough.

## **OVERVIEW**

### How should New York City's resources be distributed?

New York City's population is not distributed evenly throughout the five boroughs. Figure 1 shows that Brooklyn and Queens have the highest percentage of the population - at 31% and 27%. The Bronx and Manhattan have roughly 10% less - at 17% and 19%. Staten Island has the least amount of people - at about 6%.

New York City should distribute resources based on the need of each borough. Therefore,  the distribution of poverty indicators, rather than the population, is more important. Figure 2 demonstrates the distribution of evictions and SNAP recipients by borough. This graph shows that the Bronx and Brooklyn have between 28% and 35% of residential evictions and SNAP recipients in New York City. This is the highest percentage in all boroughs, indicating that residents of these two boroughs may have a higher need for resources. Manhattan and Queens have between 14% and 19% of residential evictions and SNAP recipients. Staten Island has the least of all with 4% of residential evictions and SNAP recipients.

### How is New York City distributing resources to help alleviate poverty?

Figure 4 depicts the distribution of resources aimed to help people experiencing poverty. These resources are defined in the legend as homebase offices, shelters, employment programs, financial empowerment centers, community centers, and SNAP centers.

- The homebase offices are locations where people who are at risk of experiencing homelessness can get resources such as temporary housing. 

- Shelters are homeless shelters where people experiencing homelessness can live.

- Employment programs are programs for youth and adults.

- Financial empowerment centers are places where people can get training and advice to better manage their finances and get out of poverty.

- Community centers are places that are open to the public for recreation. These include senior centers, family centers, and more. 

- SNAP centers are locations where people can receive benefits such as food assistance.

Figure 4 shows that Brooklyn, the Bronx, and Manhattan generally have a higher percentage of resources than Queens and Staten Island. However, this is not true for every benefit. (See Figure 5 for detailed plots on each resource.) Nonetheless, it does demonstrate that New York City is allocating resources well based on need because Figure 2 shows that Brooklyn and the Bronx have the highest need.

Figure 1

```{r overview_population_bar}

# NYC Population Bar Graph

population_bar

```

Figure 2 

```{r overview_poverty_indicator_bar}

# POVERTY INDICATOR (percent) Bar Graph

borough_indicator_bar

```

Figure 3 

```{r overview_borough_indicator_table}

# Table of POVERTY INDICATORS (count) by BOROUGH

borough_indicator_table

# Removing objects

rm(population_bar, borough_indicator_bar, borough_indicator_table)

```



Figure 4

```{r overview_resources_bar}

# NYC Resources Bar Graph

borough_resources_percent_bar

# Removing the object from the global environment

rm(borough_resources_percent_bar)

```

<details>

<summary> ***See below for detailed plots*** </summary>

Figure 5

```{r overview_resources_bar_detailed}

borough_resources_percent_bar_detailed

# Removing the object from the global environment

rm(borough_resources_percent_bar_detailed)

```

</details>


Figure 6

```{r overview_resources_table}

# Table of RESOURCES (count) by BOROUGH

borough_resources_count_table

# Removing the object from the global environment

rm(borough_resources_count_table)

```

### Statistics on Resources in New York City

Figure 7

Standard Deviation has been rounded to whole numbers.

```{r overview_five}

# NYC Resources Statistics

borough_resource_stats_table

# Removing the object from the global environment

rm(borough_resource_stats_table)

```

## **BRONX**{.tabset}

The Bronx has 22,119 residential evictions and 7,819,839 SNAP recipients (see Figure 3). This is the highest amount of residential evictions in New York City, and the second-highest amount of SNAP recipients. Brooklyn and the Bronx have similar amounts for these poverty indicators. As such, the Bronx and Brookly should have the most resources to meet this need.

The Bronx has at least 10,000 more residential evictions than Queens, Manhattan, and Staten Island. Yet, the distribution of shelters and SNAP centers is almost even between Brooklyn, the Bronx, Queens, and Manhattan (see Figure 5). The Bronx and Brooklyn should have significantly more shelters and SNAP Centers compared to Queens and Manhattan.

The Bronx and Brooklyn have the highest percentage of homebases compared to other boroughs. This will serve the Bronx community well because they have the highest amount of residential evictions.

Employment Programs are distributed evenly between the Bronx, Queens, and Manhattan, with Brooklyn having significantly more and Staten Island having significantly less. Given that the Bronx has more poverty indicators than the other boroughs besides Brooklyn (see Figure 5), it does not seem to be the best allocation of resources to meet the needs of these communities.

The distributions of financial empowerment centers and community centers have a similar shape (see Figure 5). The Bronx has significantly fewer of these resources than Brooklyn and Manhattan. The amount allocated to Brooklyn is justified by their demonstrated need (see Figure 2), but the amount allocated to Manhattan does not match their need because they have the 4th lowest amount of SNAP residents and residential evictions of all five boroughs. Once again, Brooklyn and the Bronx should have the highest amount of these resources.

Overall, it is good that most homebases are located in the Bronx and Brooklyn because they can benefit more than other boroughs from that resource. However, there should be more shelters, SNAP centers, financial empowerment centers, employment centers, and community centers, given that they have more people in need to serve.

Figure 8

```{r bronx_resources_table}

# Bronx Table

bronx_resources_table

```

Figure 9

```{r bronx_resources_bar}

# Bronx Bar Graph

bronx_resources_bar

# Add a table of stats for the Bronx.

```

## **BROOKLYN**

Brooklyn has 18,898 residential evictions and 9,745,578 SNAP recipients (see Figure 3). This is the second-highest number of residential evictions in New York City and the highest number of SNAP recipients. Brooklyn and the Bronx have similar amounts for these poverty indicators. As such, the Bronx and Brookly should have the most resources to meet this need.

As stated in my analysis of the Bronx, the distribution of shelters and SNAP centers is almost even between Brooklyn, the Bronx, Queens, and Manhattan (see Figure 5). However, given the difference in need, as represented by poverty indicators, Brooklyn and the Bronx should have significantly more shelters and SNAP Centers than Queens and Manhattan.

Brooklyn and the Bronx have the highest percentage of homebases compared to other boroughs. This will serve the Brooklyn community well because they have the second highest amount of residential evictions.

Brooklyn has significantly more employment programs than any other borough. This is an excellent example of distributing resources according to need, given that Brooklyn has some of the highest poverty indicators, along with the Bronx.

Financial empowerment centers and community centers distributions have a similar shape (see figure 5). In the graph, we can see that Brooklyn has the second highest amount after Manhattan. Given that Manhattan has the 4th lowest amount of SNAP residents and residential evictions of all five boroughs, it does not justify the need for the most financial empowerment centers. Manhattan is a central location, but it may still be inaccessible to those who live in the outer boroughs. Thus, resources should be distributed based on need, so those in need can easily access these resources.

Overall, it is good that most homebases are located in Brooklyn and the Bronx because they can benefit more than other boroughs from that resource. However, there should be more shelters, SNAP centers, financial empowerment centers, employment centers, and community centers, given that they have more people in need to serve.

Figure 10

```{r brooklyn_table}

# Brooklyn Table

brooklyn_resources_table

```

Figure 11

```{r brooklyn_resources_bar}

# Brooklyn Resources Bar Graph

brooklyn_resources_bar

```

## **MANHATTAN**

Manhattan has 8,910 evictions and 3,909,131 SNAP recipients, the second-lowest amount of all the boroughs. In contrast, Manhattan has more community centers, financial empowerment centers, and SNAP centers than other boroughs, which does not match their need according to the poverty indicators because Manhattan is among the least needy boroughs (see figure 2). 

Manhattan has an almost equal percentage of employment programs and shelters with Queens, Brooklyn, and the Bronx. Manhattan, along with Queens and Staten Island, has a homebase count significantly lower than the Bronx and Brooklyn.

Figure 12

```{r manhattan_resources_table}

# Manhattan Table

manhattan_resources_table

```

Figure 13

```{r manhattan_resources_bar}
# Manhattan Bar Graph

manhattan_resources_bar

```


## **QUEENS**

Queens has 12,653 evictions and 5,136,808 SNAP recipients, which is the third largest amount for all the boroughs. The amount of need in Queens is close to the amount in Manhattan, so the distribution of resources in these boroughs should be similar (see figure 2). 

Queens, Manhattan, and the Bronx have an equal number of SNAP centers, employment programs, and shelters. This is proportionate to the need since Queens is the third neediest borough. However, Queens has one of the lower amounts of financial empowerment centers, community centers, and homebases. In the distribution of community centers, there is a large gap between amounts in Queens compared to Manhattan, Brooklyn, and the Bronx. Since Queens has slightly more need than Manhattan, they should have a similar quantity of community centers. 

Figure 14

```{r queens_resources_table}

# Queens Table

queens_resources_table

```

Figure 15

```{r queens_resources_bar}
# Queens Bar Graph

queens_resources_bar

```

## **STATEN ISLAND**

Staten Island has 2,321 evictions and 10,84,188 SNAP recipients, the least among all five boroughs by a significant amount (see figure 2). The distribution of all the resources, aside from homebases, shows Staten Island has the least amount, which is proportionate to their demonstrated need (see figure 5). In regards to the distribution of homebases, Staten Island, Queens, and Manhattan share the least amount of homebases per borough by far. However, these distributions should be different because Staten Island has less need than Queens and Manhattan. Nevertheless, overall the distribution of the various resources in Staten Island is proportionate to the needs distribution. 

Figure 16

```{r staten_island_resources_table}

# Staten Island Table

staten_island_resources_table

```

Figure 17

```{r staten_island_resources_bar}

# Staten Island Bar Graph

staten_island_resources_bar

```

## **CONCLUSION**

The Bronx and Brooklyn have the highest need. However, surprisingly, the graphs do not show a significant increase in resources like they do in the graph on poverty indicators, with homebases being an expectation. Nevertheless, for all the resource distributions, they are at least among the top three boroughs in addition to Manhattan.

The distribution of employment programs, shelters, and SNAP centers has an almost equal percentage across the four boroughs of the Bronx, Brooklyn, Queens, and Manhattan, with somm exceptions (see statistics tab for more analysis on these resources)

Staten Island has the least need by a significant amount, which is clearly reflected in the distribution of resources as they have a tiny proportion of each resource.


## ***STATISTICS***

I will use ANOVA tables to confirm the relationships between the boroughs of Bronx, Brooklyn, Queens, and Manhattan and the resources of employment programs, shelters, and SNAP centers that were visualized in the bar graphs in figure 5.

In my conclusion, I stated that these resources appear to be almost equally distributed in the graphs (see figure 5), with some variation, such as one borough being higher. The F Statistics in the ANOVA table demonstrates the strength of a model and provides a number to compare. I will compare the strengths of models made up of one borough (using dummy variables) and one of the above resources.

### Employment programs

The F statistic for Queens and Manhattan, at 0.000, is equal. For the Bronx, it is slightly higher at 0.020. Lastly, for Brooklyn, it is significantly higher at 4.009, evident in figure 5 due to the higher bar. This indicates that the relationship between Brooklyn and Employment programs is more substantial than in other boroughs. 

```{r employment_programs_ANOVA}

bronx_employment_model <- lm(borough_bronx ~ employment_programs_count, data = borough_resources_count)

supernova(bronx_employment_model)

brooklyn_employment_model <- lm(borough_brooklyn ~ employment_programs_count, data = borough_resources_count)

supernova(brooklyn_employment_model)

queens_employment_model <- lm(borough_queens ~ employment_programs_count, data = borough_resources_count)

supernova(queens_employment_model)

manhattan_employment_model <- lm(borough_manhattan ~ employment_programs_count, data = borough_resources_count)

supernova(manhattan_employment_model)

```

### Shelters

The F statistic for Queens and the Bronx is equal at 0.066. For Manhattan, it is slightly lower at 0.035. Lastly, for Brooklyn, it is significantly higher at 1.280, which is evident in figure 5 due to the higher bar. This indicates that the relationship between Brooklyn and shelters is more robust than in other boroughs. Furthermore, when we compare the `brooklyn_employment_model` with the `brooklyn_shelter_model`, the F statistics also indicate a stronger relationship with employment programs than shelters in Brooklyn.

```{r shelters_ANOVA}

# Bronx

bronx_shelter_model <- lm(borough_bronx ~ shelters_count, data = borough_resources_count)

supernova(bronx_shelter_model)

# Brooklyn

brooklyn_shelter_model <- lm(borough_brooklyn ~ shelters_count, data = borough_resources_count)

supernova(brooklyn_shelter_model)

# Queens

queens_shelter_model <- lm(borough_queens ~ shelters_count, data = borough_resources_count)

supernova(queens_shelter_model)

# Manhattan

manhattan_shelter_model <- lm(borough_manhattan ~ shelters_count, data = borough_resources_count)

supernova(manhattan_shelter_model)

```

### SNAP centers

The F statistic for Brooklyn, Bronx, and Queens, at 0.032, is equal. However, for Manhattan, it is significantly higher at 1.800, evident in figure 5 due to the higher bar. This indicates that the relationship between Manhattan and SNAP centers is more substantial than in other boroughs.

```{r SNAP_centers_ANOVA}

# Bronx

bronx_snap_center_model <- lm(borough_bronx ~ snap_centers_count, data = borough_resources_count)

supernova(bronx_snap_center_model)

# Brooklyn

brooklyn_snap_center_model <- lm(borough_brooklyn ~ snap_centers_count, data = borough_resources_count)

supernova(brooklyn_snap_center_model)

# Queens

queens_snap_center_model <- lm(borough_queens ~ snap_centers_count, data = borough_resources_count)

supernova(queens_snap_center_model)

# Manhattan

manhattan_snap_center_model <- lm(borough_manhattan ~ snap_centers_count, data = borough_resources_count)

supernova(manhattan_snap_center_model)

```

